\Huge \textcolor{titleTextColor}{ \textsf{\textbf{ \fontsize{85}{60}\selectfont GRASS GIS: a peer-reviewed scientific platform and future research repository }} }

### 2015-03-31 grassposter/2015_EGU_G7_PeerReview_SciPlatform/poster.tex
Updated Introduction with Peter's text
git-svn-id: https://svn.osgeo.org/grass/grass-promo@64954 15284696-431f-4ddb-bdfa-cd5b030d7da7

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Introduction}}
{
\setlength{\parskip}{0.3ex}

Over the last decades, GIS has become a key driver in geospatial science, research and application.
GIS software which is licensed under a free and open source software (FOSS) licence
is more than just a mere tool for spatial analysis.

GRASS GIS (Neteler et al., 2012 \cite{neteler2012grass}), a free and open source GIS,
is used by many scientists directly or as a backend in other projects
such as R or QGIS to perform geoprocessing tasks.

Thanks to the user and developer community, submitted code is evaluated
in different fields of application beyond
the field of expertise of the original authors, and different scales of magnitude
for the data to be processed.
This exceeds the established review process for scientific writing in a given journal
or a data publication in a defined field of science.

Immediate access to software repository enables instant quality checking
of the current software version both by continuous automated tests (Petras, 2014 \cite{Petras2014}),
and code review by human experts.

New scientific algorithms can be developed against the reviewed functionalities
already provided by the GRASS GIS codebase.
This avoids unnecessary overhead by re-implementation,
ensures quality by use of trusted components and allows reuse and long term preservation
within the project software repository:
Integrating scientific algorithms into GRASS GIS helps to preserve reproducibility
of scientific results over time as the original author designed it
(Rocchini \& Neteler, 2012 \cite{rocchini2012let}).
}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Introduction}}
{
\setlength{\parskip}{0.3ex}
\normalsize
Geographical Information Systems (GIS) are widely used for the management of both natural ressources and in more general terms all kind of information with a spatial dimension associated to it. Over the last decades, GIS has become a key driver in geospatial science, research and application.

GIS software which is licensed under a free and open source software (FOSS) licence is more than just a mere tool for spatial analysis. FOSS GIS projects are maintained and continuously extended by international communities of volunteers, based on freely publicly available code base. Immediate access to this software repository enables instant quality checking of the current software version both by continuous automated tests, and code review by human experts. FOSS GIS project communities include members from many fields of Science and Industry, which leads to a many-faceted review process for incoming software submissions to the project code-base. Submitted code is evaluated in different software environments beyond the one originally used for development, in different fields of application beyond the field of expertise of the original authors, and different scales of magnitude for the data to be processed. This exceeds the established review process for scientific writing in a given journal or a data publication in a defined field of Science. FOSS GIS projects can serve as collaborative peer-reviewed environments to conduct verifiable science in a trusted and transparent environment.

GRASS GIS (Neteler et al., 2012 \cite{neteler2012grass}), a free and open source GIS, is used by many scientists directly or as a backend in other projects such as R or QGIS to perform geoprocessing tasks. Thus, a large number of current scientific geospatial computations already rely on quality and correct functionality of the software, as a means to assure the replicability of the scientific results. In addition, GRASS GIS provides added-value for new and ongoing research as a stable and reliable platform: New scientific algorithms can be developed against the reviewed functionalities already provided by the GRASS GIS codebase. This avoids unnecessary overhead by re-implementation, ensures quality by use of trusted components and allows reuse and long term preservation within the project software repository:   Integrating scientific algorithms into GRASS GIS helps to preserve reproducibility of scientific results over time as the original author designed it (Rocchini & Neteler, 2012 \cite{rocchini2012let}).

Moreover, subsequent improvements are tracked in the source code versioning system and are immediately available to the public (Petras, 2014 \cite{Petras2014}). Thus, GRASS GIS acts as a repository of scientific peer-reviewed code and algorithm/knowledge hub for future generation of scientists.
